Tight Bounds for Dynamic Balanced Graph Partitioning

### 2020-05-10 paper/disc20clustering.tex
Self-review

\begin{table*}
	\centering
	\renewcommand{\arraystretch}{1.5}
	\begin{tabular}{>{\centering\arraybackslash}p{4.5cm}|>{\centering\arraybackslash}p{4.5cm}>{\centering\arraybackslash}p{4.5cm}}
		\rowcolor{gray!50}
		\textbf{Variant} & \textbf{ Lower bound} &\textbf{Upper bound}\\ \hline 
		\textbf{$k=2$}& 3\hspace{0.3cm}\cite{repartition-disc} & 6\hspace{0.3cm}(\S \ref{sec:k2}) \\ 
		\rowcolor{gray!25}
		\textbf{$k=3$}&  $\Omega(\ell)$ \hspace{0.3cm}(\S \ref{sec:lowerbound})& $O(\ell) $\hspace{0.3cm}(\S \ref{sec:k3})\\
		$k > 3$ & $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S  \ref{sec:lowerbound})&$O(k^2 \cdot \ell^2)$\hspace{0.1cm} \cite{repartition-disc} \\
		\rowcolor{gray!25}
		Learning model & $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S  \ref{sec:lowerbound})&$O(k \cdot \ell)$\hspace{0.3cm} (\S \ref{sec:ppl}) \\
	\end{tabular}
	\caption{Overview of known results and the results provided in this paper. By $\Rep$ we denote the cost of a single component repartition, and we elaborate on this in Section~\ref{sec:k3}. The first three rows show the results for the general partitioning model, except for the last row that shows the results for the learning model for arbitrary $k$ and $\ell$.
	}
	\label{tab:overview}
	\vspace{-7mm}
\end{table*}

@@

\begin{theorem}
	\TAlg is $(54\ell)$-competitive for $k=3$.
\end{theorem}
\begin{proof}
	Fix a completed phase, and consider the state of \TAlg's counters at the end of it.
	We consider the incomplete phase later in this proof.

@@

	Finally, we bound the competitive ratio.
	From the bound on \OPT's cost we have $|\sigma_E| \geq 2(\OPT(\sigma_E)+\OPT(M)) + 4\alpha \cdot \ell$.
	Let $\xi := \OPT(\sigma_E) + \OPT(M)$.
	The competitive ratio is

	\[
		\frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{11\alpha \cdot \ell + |\sigma_E|}{\alpha/2 + \xi/2} \leq \frac{27\alpha\cdot\ell + 2\cdot \xi}{\alpha + \xi} \leq 27 \ell.
	\]

@@

	It remains to consider the last, unfinished phase.
	First, consider the case, where the unfinished phase is also the first one.
	Then, we cannot charge $\OPT$ due to the inability to partition the components.
	Instead, we use the fact that \TAlg and $\OPT$ started with the same initial configuration.
	If the input finished before the first $\alpha$ external requests, then \TAlg is $1$-competitive.
	If there were at least $\alpha$ external requests, then $\OPT$ either paid $\alpha$ for serving them remotely, or paid $\alpha$ for a migration; for the cost of \TAlg we follow the analysis of the unsaturated pairs.
	Second, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and follow the analysis regarding the unsaturated requests.
	This way, the competitive ratio increases at most twofold in comparison to a finished phase, and the competitive ratio is $\TAlg(\sigma)/\OPT(\sigma) \leq 54\ell$.
\end{proof}

>>>>

\begin{table*}
	\centering
	\renewcommand{\arraystretch}{1.5}
	\begin{tabular}{>{\centering\arraybackslash}p{4.5cm}|>{\centering\arraybackslash}p{4.5cm}>{\centering\arraybackslash}p{4.5cm}}
		\rowcolor{gray!50}
		\textbf{Variant} & \textbf{ Lower bound} &\textbf{Upper bound}\\ \hline 
		\textbf{$k=2$}& 3\hspace{0.3cm}\cite{repartition-disc} & 6\hspace{0.3cm}(\S \ref{sec:k2}) \\ 
		\rowcolor{gray!25}
		\textbf{$k=3$}&  $\Omega(\ell)$ \hspace{0.3cm}(\S \ref{sec:lowerbound})& $O(\ell) $\hspace{0.3cm}(\S \ref{sec:k3})\\
		$k > 3$ & $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S  \ref{sec:lowerbound})&$O(k^2 \cdot \ell^2)$\hspace{0.1cm} \cite{repartition-disc} \\
		\rowcolor{gray!25}
		Learning model & $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S  \ref{sec:lowerbound})&$O(k \cdot \ell)$\hspace{0.1cm} (\S \ref{sec:ppl}) \\
	\end{tabular}
	\caption{Overview of known results and the results provided in this paper. By $\Rep$ we denote the cost of a single component repartition, and we elaborate on this in Section~\ref{sec:k3}. The first three rows show the results for the general partitioning model, except for the last row that shows the results for the learning model for arbitrary $k$ and $\ell$.
	}
	\label{tab:overview}
	\vspace{-7mm}
\end{table*}

@@

\begin{theorem}
	\TAlg is $(60\ell)$-competitive for $k=3$.
\end{theorem}
\begin{proof}
	Fix a completed phase, and consider the state of \TAlg's counters at the end of it.
	We consider the incomplete phase later in this proof.

@@

	Finally, we bound the competitive ratio.
	From the bound on \OPT's cost we have $|\sigma_E| \leq 2(\OPT(\sigma_E)+\OPT(M)) + 4\alpha \cdot \ell$.
	Let $\xi := \OPT(\sigma_E) + \OPT(M)$.
	The competitive ratio is

	\[
		\frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{11\alpha \cdot \ell + |\sigma_E|}{\alpha/2 + \xi/2} \leq \frac{30\alpha\cdot\ell + 4\cdot \xi}{\alpha + \xi} \leq 30 \ell.
	\]

@@

	It remains to consider the last, unfinished phase.
	First, consider the case, where the unfinished phase is also the first one.
	Then, we cannot charge $\OPT$ due to the inability to partition the components.
	Instead, we use the fact that \TAlg and $\OPT$ started with the same initial configuration.
	If the input finished before the first $\alpha$ external requests, then \TAlg is $1$-competitive.
	If there were at least $\alpha$ external requests, then $\OPT$ either paid $\alpha$ for serving them remotely, or paid $\alpha$ for a migration; for the cost of \TAlg we follow the analysis of the unsaturated pairs.
	Second, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and follow the analysis regarding the unsaturated requests.
	This way, the competitive ratio increases at most twofold in comparison to a finished phase, and the competitive ratio is $\TAlg(\sigma)/\OPT(\sigma) \leq 60\ell$.
\end{proof}
